DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-02-01 paper/comparison.tex
revision

In this section, we compare three computing platforms---the Raspberry
Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
Jetson
TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
the point of view of supporting end-to-end deep learning
based autonomous vehicles. 
Table~\ref{tbl:platforms} shows architectural features of the three
platforms~\footnote{The GPUs of the Raspberry Pi 3 and Intel
  Up are not used in evaluation due to the lack of software (TensorFlow)
support. Also, the two Denver cores in the Tegra TX2 are not used in
evaluation due to TensorFlow issues.}.

>>>>

In this section, we compare three computing platforms---the Raspberry
Pi 3, the Intel UP~\footnote{http://www.up-board.org/up/} and NVIDIA
Jetson
TX2~\footnote{http://www.nvidia.com/object/embedded-systems-dev-kits-modules.html}---from
the point of view of supporting end-to-end deep learning
based autonomous vehicles. 
Table~\ref{tbl:platforms} shows architectural features of the three
platforms~\footnote{The GPU of Intel UP and the two Denver cores in the
  Tegra TX2 are not used in evaluation due to TensorFlow issues.}.

### 2018-02-01 paper/evaluation.tex
revision

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting for about 81\% of
the execution time.
Second, and more importantly, we also find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at a 30 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).

In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs also at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that the low-cost
Raspberry Pi 3 can achieve similar real-time control performance is
surprising.

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
the control loop as we vary the number of cores used by
TensorFlow. As expected, as we assign more cores, the average execution
time decreases---from 46.30 ms on a single core to 22.86 ms on four
cores (over a 100\% improvement). However, the improvement is far from an ideal
linear scaling. In particular, from 3 cores to 4 cores, the
improvement is mere 2.80 ms, or 12\%. In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform. We do not know whether it is due to the limitations of
TensorFlow's multicore implementation or if it's the model's inherent
characteristics.

The poor scalability opens up the possibility of consolidating
multiple different tasks or different neural network models rather
than allocating all cores for a single neural network model. For
example, it is conceivable to use four cameras and four different
neural network models, each of which is trained separately and
executed on a single dedicated core. Assuming we use the same network
architecture for all models, then one might expect to achieve up to
20 Hz using one core (given 1 core can deliver 46 ms average
execution time). In the next experiment, we investigate the
feasibility of such a scenario.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/l2missrate_vs_modelcnt}
  \caption{L2 cache miss rates of different neural network and core
    assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
  \label{fig:l2missrate-vs-modelcnt}
\end{figure}

These increases in inference times in the co-scheduled scenarios are
expected and likely caused by contention in the shared hardware
resources, such as the shared L2 cache and/or the DRAM controller.
To further analyze the source of contention, we use hardware
performance counters of the processor. Specifically, we measure L2
miss rates of the DNN models first in isolation and then after
co-scheduling other models. If the shared L2 cache is the primary
source of inteference, then the measured L2 miss rates will
increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
As can be see in the figure, L2 miss rates increase as more models are 
co-scheduled, but the increase is less than expected.
This suggests that the shared L2 cache is not the main bottleneck that 
causes execution time increases. In other words, DNN models don't appear 
to be significantly affected by the shared L2 cache space. 
% L2 partitioning is probably not going to be useful.
Instead, we hypothesize that it is likely caused by the memory
controller---the only other major shared hardware source---where
memory requests from different CPU cores contend, which would result
in increased memory access latency. While some Intel processors
provide incore hardware counters that can measure average memory
access latency~\cite{ye2016maracas}, we were not able to identify
whether such hardware counters exist in the BCM2837 processor of
Raspberry Pi 3 due to the lack of documentation. Instead, in the next
experiment, we use memory intensive synthetic benchmarks to test the
hypothesis.

@@

In order to determine how contended DRAM requests affect the DNN
inference timing of the DeepPicar, we use a synthetic memory
intensive benchmark from the IsolBench suite~\cite{Valsan2016}.

We run a single DNN model on one core, and
co-schedule an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384}}, on the remaining idle cores.

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 11.6X---even though the DNN model is running on a
dedicated core (Core 0). On the other hand, the DNN model's L2
cache-miss rates do not increase as much. This suggests that
the DNN model's execution increase cannot be fully explained by
increases in L2 cache-misses. Instead, as we hypothesized in the previous
experiment, the increased memory pressure from the co-scheduled memory
intensive benchmarks is likely the primary cause of the DNN model's execution
time increase. Therefore, we conclude that DeepPicar's DNN model is
more senstive to DRAM access latency than L2 cache space.

This observation suggests that shared cache partitioning
techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
solutions for DeepPicar, as its AI workload is more sensitive to memory
performance. Instead, memory controller focused isolation solutions,
either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
may be more important. Although our observation is made on a single
hardware platform running on a single DNN workload, we suspect that
many AI workloads may exhibit similar characteristics.

\subsection{Effect of Cache Partitioning}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{figs/cache-mapping}
  \caption{A memory mapping of the cache sets for the Broadcom 
BCM2837 processor used in the Raspberry Pi 3 Model B.}
  \label{fig:cache-mapping}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_multicore}
  \caption{Average control loop execution time vs. \# % of L2 cache
  space available.}
  \label{fig:palloc_multicore}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_multicore_l2missrate}
  \caption{L2 miss rate impact of limiting the amount of L2 cache space
  available to the DNN.}
  \label{fig:palloc_multicore_l2missrate}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_multimodel}
  \caption{Timing impact of co-scheduling multiple DNNs when cache 
partitioning is enabled.}
  \label{fig:palloc_multimodel}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Timing impact of co-scheduling memory intensive co-runners 
when cache partitioning is enabled.}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_bus}
  \caption{Bus access impact of co-scheduling memory intensive co-runners 
when cache partitioning is enabled.}
  \label{fig:palloc_bandwidth_bus}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.7\textwidth]{figs/palloc_bandwidth_l2missrate}
  \caption{L2 miss rate impact of co-scheduling memory intensive co-
runners when cache partitioning is enabled.}
  \label{fig:palloc_bandwidth_l2missrate}
\end{figure}

In order to determine if the DNN is mostly affected by the DRAM 
controller, we partition the L2 cache of the Pi3 and reperform the 
same experiments to see if there are any noticable changes. For the 
partition, we employ PALLOC~\cite{yun2014rtas}, a color-based page 
allocator that works on the kernel level. For the bit mask, we select 
bits 12, 13 and 14 as they can be used to access the L2 cache,
as can be seen by Figure \ref{fig:cache-mapping}. This results in 
2\textsuperscript{3} = 8 colors, which we then assign to the Pi3's 
physical cores such that each core has two unique colors (colors 0 
and 1 are assigned to core 0, colors 2 and 3 are assigned to core 1, 
etc.). In the case of bit 12, since we use 2 colors for each partition,
the L1 Data cache is not partitioned in our experiments.

If performance improves as a result of partitioning the shared cache 
then we know that the DNN, to some extent, relies on shared cache 
performance in addition to shared DRAM performance. On the other 
hand, if there is no improvement in performance, then it can be 
observed that the shared memory is more critical to the performance 
of the DNN.

We run the DNN model with different amounts of L2 cache being
available based on the number of colors being used. For example, if two 
colors are used, then 25\% of the L2 cache is available, whereas if all 
8 colors are used, 100\% of the L2 cache is available. Particularly, we 
test the performance of the model when it utilizes a single core and
all four of the available cores. The results can be seen in 
\ref{fig:palloc_multicore}. As expected, the amount of L2 cache 
available to the model does not result in any significant changes in 
performance. This can also be seen by the L2 miss rates of the model 
seen in \ref{fig:palloc_multicore_l2missrate}. As more cache space 
becomes avaiable, the percentage of L2 misses decreases, which doesn't 
affect the real-time performance of the DNN. As such, we find that a 
single model is not sensitive to the shared L2 cache.

We also co-schedule multiple models to see how cache partitioning affects
interference between them. Each model uses an equal number of colors, and
are thus given equal amounts of L2 cache space, and there is no overlap
in the colors used, so no L2 cache is shared between models. As can be 
seen in \ref{fig:palloc_multimodel}, the performance of the models 
remains the same as when no cache partitioning was employed. Based of 
these results, we find that contention for the L2 cache is not the main
source of interference between multiple DNN models.

Finally, we run a single model alongside memory intensive co-runners 
again, but ensure that each task is given equal amounts of L2 cache space.
For this, we assign two colors to each task, regardless of the number of
tasks being scheduled. As a result, the entire cache is not utilized in
experiments with less than four tasks. The results can be seen in Figure
\ref{fig:palloc_bandwidth_exectime}. With zero or one co-runners, the DNN
performance doesnt't change relative to when the cache wasn't partitioned.
However, we find that with two or more co-runners, DNN performance becomes
worse. In the case of two co-runners, average control loop time increases
by 66.31 ms, or 26\%, and with three co-runners, the average time
is 265.64 ms, of 49\%, greater.
At the same time, though, the L2 miss rates of the model don't correlate 
to the timing increases seen, as is shown in 
\ref{fig:palloc_bandwidth_l2missrate}. Instead, we find that the number
of bus accesses increases when cache partitions are used, and they mirror
the timing increases of the model. This can be seen in Figure 
\ref{fig:palloc_bandwidth_bus}. These additional bus accesses are
most likely caused by the L2 prefetcher, as it can generate additional bus
accesses that aren't counted as L2 misses. As a result, we find that cache
partitioning is detrimental to the performance of the DNN so long as the 
L2 prefetcher is enabled (we have not yet found a way to disable it on
the Pi 3). Furthermore, no cache sensitivity was displayed by the model,
thus showing that the performance of the L2 cache is not vital for DNN 
performance when co-runners are introduced.

By partitioning the shared L2 cache, we find that no noticable 
improvements are gained and that performance remains consistent, or 
decreases, depending on the circumstances. In all experiements, the DNN
shows no sensitivity to the shared L2 cache. As a result, we conclude 
that cache partitioning is not an effective isolation mechanism, and that 
the performance of the shared DRAM controller is of greater importance 
to the real-time efficiacy of the DeepPicar.

@@

Lastly, we find that partitioning the shared L2 cache provides 
practically no benefits to the real-time performance of the DNN. To the 
contrary, performance is constant regardless of the amount of cache 
space used. We also find that performance becomes worse when memory
intensive co-runners are introduced as they generate more memory 
contention when partitioning is enabled. Furthermore, we 
find that performance remains consistent when multiple models are run 
simultaneously. As such, we conclude that the shared L2 cache is not 
essential to the DNN's real-time performance.

>>>>

First, as expected, we find that the inference operation
dominates the control loop execution time, accounting for about 81\% of
the execution time.

Second, we find that the measured average
execution time of a single control loop is 22.87 ms, or 43.7 Hz and
the 99 percentile time is 23.38 ms.
This means that the DeepPicar can operate
at a up to 40 Hz control frequency in real-time using only the on-board
Raspberry Pi 3 computing platform, as no remote computing resources 
were necessary. We consider these results surprising given the complexity
of the deep neural network, and the fact that the inference operation
performed by TensorFlow only utilizes the CPU cores of the
Raspberry Pi 3 (its GPU is not supported by Tensorflow).
In comparison, NVIDIA's DAVE-2 system, which has the exact same neural
network architecture, reportedly runs at 30 Hz~\cite{Bojarski2016}. 
Although we believe it was not
limited by their computing platform (we will experimentally compare
performance differences among multiple embedded computing platforms,
including NVIDIA's Jetson TX2, later in
Section~\ref{sec:comparison}), the fact that a low-cost
Raspberry Pi 3 can achieve comparable real-time control performance is
surprising.

Lastly, we find that the control loop execution timing is highly
predictable and shows very little variance over different input image
frames. This is because the amount of computation needed to perform
a DNN inferencing operation is fixed at the DNN architecture design
time and does not change at runtime over different inputs (i.e.,
different image frames). This predictable timing behavior is a highly
desirable property for real-time systems, making DNN inferencing an
attractive real-time workload.

@@

Figure~\ref{fig:perf-vs-corecnt} shows the average execution time of
the control loop as we vary the number of cores used by
TensorFlow. As expected, as we assign more cores, the average execution
time decreases---from 46.30 ms on a single core to 22.86 ms on four
cores (over a 100\% improvement). However, the improvement is far from an ideal
linear scaling. In particular, from 3 cores to 4 cores, the
improvement is mere 2.80 ms, or 12\%. In short, we find that the
scalability of DeepPicar's deep neural network is not ideal on the
platform.

As noted in~\cite{NVIDIA2015}, DNN inferencing is inherently more
difficult to parallelize than training because the easiest
parallelization option, batching (i.e., processing multiple images in
parallel), is not available or is limited. Specifically, in DeepPicar,
only one image frame, obtained from the camera, can be processed at a
time. Thus, more fine-grained algorithmic parallelization is needed to
improve inference performance~\cite{NVIDIA2015}, which generally does
not scale very well.

On the other hand, the poor scalability opens up the possibility of
consolidating multiple different tasks or different neural network
models rather than allocating all cores for a single neural network
model.
For example, it is conceivable to use four cameras and four different
neural network models, each of which is trained separately for
different purpose and executed on a single dedicated core.
Assuming we use the same network
architecture for all models, then one might expect to achieve up to
20 Hz using one core (given 1 core can deliver 46 ms average
execution time).
In the next experiment, we investigate the
feasibility of such a scenario.

@@

These increases in inference times in the co-scheduled scenarios are
expected and are caused by contention in the shared hardware
resources, such as the shared L2 cache and/or the DRAM controller.

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.7\textwidth]{figs/l2missrate_vs_modelcnt}
%%   \caption{L2 cache miss rates of different neural network and core
%%     assignments. X-axis is the same as Figure~\ref{fig:perf-vs-modelcnt}.} 
%%   \label{fig:l2missrate-vs-modelcnt}
%% \end{figure}

%% To further analyze the source of contention, we use hardware
%% performance counters of the processor. Specifically, we measure L2
%% miss rates of the DNN models first in isolation and then after
%% co-scheduling other models. If the shared L2 cache is the primary
%% source of inteference, then the measured L2 miss rates will
%% increase. Figure~\ref{fig:l2missrate-vs-modelcnt} shows the results.
%% As can be see in the figure, L2 miss rates increase as more models are 
%% co-scheduled, but the increase is less than expected.
%% This suggests that the shared L2 cache is not the main bottleneck that 
%% causes execution time increases. In other words, DNN models don't appear 
%% to be significantly affected by the shared L2 cache space. 
%% % L2 partitioning is probably not going to be useful.
%% Instead, we hypothesize that it is likely caused by the memory
%% controller---the only other major shared hardware source---where
%% memory requests from different CPU cores contend, which would result
%% in increased memory access latency. While some Intel processors
%% provide incore hardware counters that can measure average memory
%% access latency~\cite{ye2016maracas}, we were not able to identify
%% whether such hardware counters exist in the BCM2837 processor of
%% Raspberry Pi 3 due to the lack of documentation. Instead, in the next
%% experiment, we use memory intensive synthetic benchmarks to test the
%% hypothesis.

@@

In this experiment, we investigate the impact of contended DRAM
requests to the DNN inference timing of DeepPicar.
For the experiment, we use a synthetic memory
intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
We run a single DNN model on one core, and
co-schedule an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384}}, on the remaining idle cores.

@@

Figure~\ref{fig:} shows the normalized execution time and L2 miss-rate
of the DNN model running on the Core 0 as a function of the number of
co-scheduled memory intensive synthetic benchmarks. First, as we
increase the number of co-runners, the DNN model's execution times are
increased---by up to 11.6X---even though the DNN model is running on a
dedicated core (Core 0). On the other hand, the DNN model's L2
cache-miss rates do not increase as much.
In other words, the DNN model's execution increase is not fully
explained by increased contention in the L2 cache space.
Instead, the high memory bandwidth pressure from the co-scheduled
memory intensive benchmarks is likely the primary cause.

%% Instead, as we hypothesized in the previous
%% experiment, the increased memory pressure from the co-scheduled memory
%% intensive benchmarks is likely the primary cause of the DNN model's execution
%% time increase. Therefore, we conclude that DeepPicar's DNN model is
%% more senstive to DRAM access latency than L2 cache space.

This observation suggests that shared cache partitioning
techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
solutions for DeepPicar's DNN processing workload, as cache
partitioning does not provide memory performance guarantees.

%% Instead, memory controller focused isolation solutions,
%% either hardware or software-based ones (e.g.,~\cite{Guo2017,Yun2013}),
%% may be more important. Although our observation is made on a single
%% hardware platform running on a single DNN workload, we suspect that
%% many AI workloads may exhibit similar characteristics.

% \input{evaluation-cachepart}

@@

%% Lastly, we find that partitioning the shared L2 cache provides 
%% practically no benefits to the real-time performance of the DNN. To the 
%% contrary, performance is constant regardless of the amount of cache 
%% space used. We also find that performance becomes worse when memory
%% intensive co-runners are introduced as they generate more memory 
%% contention when partitioning is enabled. Furthermore, we 
%% find that performance remains consistent when multiple models are run 
%% simultaneously. As such, we conclude that the shared L2 cache is not 
%% essential to the DNN's real-time performance.

### 2018-02-01 paper/overview.tex
Changes to paper
video links
revision

\begin{figure}[t]
  \centering
  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to replicate the
NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, the Raspberry Pi 3. Because the Pi 3's computing
performance is much lower than that of the DRIVE PX platform used in
DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
frequency---locally on the Pi 3 platform, although training of the 
network (back-propagation for weight updates) can be done offline and 
remotely using a desktop computer.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{|c|r|}
    \hline
    Item                    & Cost (\$) \\
    \hline
    Raspberry Pi 3 Model B  & 35 \\
    New Bright 1:24 scale RC car       & 10 \\
    Playstation Eye camera  &  7 \\
    Pololu DRV8835 motor hat&  8 \\
    External battery pack \& misc.   & 10 \\
    \hline
    Total                   & 70 \\
    \hline
  \end{tabular}
  \caption{DeepPicar's bill of materials (BOM)}
  \label{tbl:carbom}
\end{table}

For the neural network architecture, we adopt a TensorFlow version of
NVIDIA DAVE-2's convolutional neural network (CNN), published by Dr.
Fridman at  MIT~\footnote{https://github.com/lexfridman/deeptesla}. The 
main difference is that we do not utilize a normalization layer, and 
instead initialize the weights using a Xavier initialization. As
in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture, which
is comprised of 8 layers, 250K parameters, and about 27 million
connections.

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 8 layers (5
    convolutional, 3 fully-connected layers), 27 million connections,
    250K parameters. The architecture is almost identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=.65\textwidth]{figs/track_new2}
  \caption{One of the custom tracks used for training/testing.}
  \label{fig:track}
\end{figure}

\begin{figure}[t]
   \lstset{language=python,
           basicstyle=\ttfamily\small,
           keywordstyle=\color{blue}\ttfamily,
           stringstyle=\color{red}\ttfamily,
           commentstyle=\color{green}\ttfamily
          }  
  \lstinputlisting[language=python]{control.py}
  \caption{Control loop}
  \label{fig:controlloop}
\end{figure}

@@

@@

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we trained.

>>>>

In this section, we provide an overview of our DeepPicar platform. In
developing DeepPicar, one of our primary goals is to faithfully
replicate NVIDIA's DAVE-2 system on a smaller scale with using a low cost
multicore platform, the Raspberry Pi 3. Because the Raspberry Pi 3's
computing performance is much lower than that of the DRIVE PX platform
used in DAVE-2, we are interested in if, and how, we can process
computationally expensive neural network operations in
real-time. Specifically, inferencing (forward pass processing)
operations must be completed within each control period
duration---e.g., a WCET of 33.$\overline{\mbox{33}}$ ms for 30Hz control 
frequency---locally on the Pi 3 platform, although training of the 
network (back-propagation for weight updates) can be done offline and 
remotely using a desktop computer.

\begin{figure}[h]
  \centering
  \includegraphics[width=.55\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|r|}
    \hline
    Item                    & Cost (\$) \\
    \hline
    Raspberry Pi 3 Model B  & 35 \\
    New Bright 1:24 scale RC car       & 10 \\
    Playstation Eye camera  &  7 \\
    Pololu DRV8835 motor hat&  8 \\
    External battery pack \& misc.   & 10 \\
    \hline
    Total                   & 70 \\
    \hline
  \end{tabular}
  \caption{DeepPicar's bill of materials (BOM)}
  \label{tbl:carbom}
\end{table}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/architecture}
  \caption{DeepPicar's neural network architecture: 8 layers (5
    convolutional, 3 fully-connected layers), 27 million connections,
    250K parameters. The architecture is almost identical to the one 
	used in NVIDIA's real self-driving car~\cite{Bojarski2016}.}
  \label{fig:architecture}
\end{figure}

For the neural network architecture, we implement
NVIDIA DAVE-2's convolutional neural network (CNN) using TensorFlow,
based on ~\cite{deeptesla}. Note, however, that the CNN model
in~\cite{deeptesla} was considerably bigger than NVIDIA's CNN as it contained an
additional fully-connected layer of approximately 1.3M additional
parameters. We eliminated the additional layer to faithfully recreate
the NVIDIA's original CNN model.
%% The main difference is that we do not utilize a normalization layer, and 
%% instead initialize the weights using a Xavier initialization. 
As in DAVE-2, the CNN takes a raw color image (200x66 RGB pixels)
as input and produces a single steering angle value as
output. Figure~\ref{fig:architecture} shows the network architecture
used in this paper, which
is comprised of 8 layers, 250K parameters, and about 27 million
connections. It is identical to NVIDIA DAVE-2's CNN except that we do
not apply a normalization layer as it does not include trainable
parameters and its computational demand is minimal.

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth]{figs/track_new2}
  \caption{One of the custom tracks used for training/testing.}
  \label{fig:track}
\end{figure}

@@

\begin{figure}[h]
   \lstset{language=python,
           basicstyle=\ttfamily\small,
           keywordstyle=\color{blue}\ttfamily,
           stringstyle=\color{red}\ttfamily,
           commentstyle=\color{green}\ttfamily
          }  
  \lstinputlisting[language=python]{control.py}
  \caption{Control loop}
  \label{fig:controlloop}
\end{figure}

@@

Despite these issues, the trained models were still able to
achieve a reasonable degree of accuracy, successfully navigating
several different tracks we trained. In our testing, we found that the
DeepPicar could remain on a track for over 10
minutes~\footnote{A collection of self-driving videos can be found at: \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}}.
